the implications of local interactions on the resilience of arid rangelands

### 2016-06-24 manuscript/manuscript.Rmd
update manusript draft: add code for livestock package; add structure for analysis and discussion

library(livestock)

@@

- The conceptualisation of rangelands as systems that can experience abrupt shifts between alternative stable states has been happening since the 70s (Noy Meir, Holling, May). 
- The responsible mechanisms are feedbacks on the landscape scale, i.e. dependent on the 'global' density of vegetation and livestock, as well as global environmental parameters, such as aridity.
- these studies had hughe implications for rangeland management and were well received in applied, fundamental and theoretical research. 
   - applied concepts such as state and transition models, i.e. descriptions of landscapes as systems with multiple potential states and probabilities and conditions for the transition between those. 
   - fundamental research addressed the existence of alternative stable states in observational and experimental studies (examples in rangelands?? arid shrublands? grasslands?). It relates to the study of successional states. 
   - theoretical studies were addressing mechanisms and behaviour of alternative stable states along gradients of pressure, where the rangeland system provided a good example. (Rietkerk, van de Koppel, Scheffer, ...) Most recently, the study of social-ecological systems addressed the catastrophic behaviour of ecosystems in the context of anthropogenic activities. (Carpenter, Schlüter)

@@

```{r, echo = FALSE, fig.cap= "Noy Meir graphs a) M vs G function, b) bifurcation along livestock density"}
plot_meanfield(livestock)

@@

Figure 3 | 3D graph of growth vs. mortality

@@

- To investigate the role of local interactions between plants, we implement these mechanisms into the model. This is complemented with figures.

>>>>

library(livestock)
library(knitr)
library(rgl)
library(foreach)
library(doSNOW)

knit_hooks$set(rgl = hook_rgl)

workerlist <- c(rep("localhost", times = 3))
cl <- makeSOCKcluster(workerlist)
 registerDoSNOW(cl)

@@

- The conceptualisation of rangelands as systems that can experience abrupt shifts between alternative stable states has been develloped since the 70s (Noy Meir, Holling, May). 
- The responsible mechanisms are feedbacks on the landscape scale, i.e. dependent on the 'global' density of vegetation and livestock, as well as global environmental parameters, such as aridity.
- these studies had hughe implications for rangeland management and were well received in applied, fundamental and theoretical research. 
   - applied concepts such as state and transition models, i.e. descriptions of landscapes as systems with multiple potential states and probabilities and conditions for the transition between those. 
   - fundamental research addressed the existence of alternative stable states in observational and experimental studies (examples in rangelands?? arid shrublands? grasslands?). It relates to the study of successional states. 
   - theoretical studies were addressing mechanisms and behaviour of alternative stable states along gradients of pressure, where the rangeland system provided a good example. (Rietkerk, van de Koppel, Scheffer, ...) Most recently, the study of social-ecological systems addressed the catastrophic behaviour of ecosystems in the context of anthropogenic activities. (Carpenter, Schlüter)

@@

```{r, echo = FALSE, fig.cap="Noy Meir graphs a) M vs G function, b) bifurcation along livestock density"}
par(mfrow = c(1,2))
p <- set_parms(livestock$defparms, set = list(b = 0.5))
plot_meanfield(livestock, new = TRUE, parms = p)
plot_bifurcation(livestock, over = "L", xrange = c(0,15), res = 51, parms = p)
```

@@

```{r, echo = FALSE, fig.cap = "3D graph of growth vs. mortality" , rgl=TRUE}
p <- set_parms(livestock$defparms, set = list(b = 0.3, c = 0.2, f = 0.9, p = 0.99, alpha = 0.2))
plot_pairapproximation3D(livestock, parms = p)
```

Figure 3 |

@@

- To investigate the role of local interactions between plants, we subsequently implement these mechanisms into the model. We investigate their impact on the ecosystem state along gradients of livestock pressure (i.e. bifurcation diagrams) under a given environmental quality, that leads to the expression of catastrophic shifts. We discuss positive and negative effects on plant growth and mortality.

## local facilitation

- local facilitation is the positive local effect of plants on the growth of neighboring plants. Plants in arid environments provide shade and increase water retention, as well as improving soil organic matter by adding and capturing litter. This leads to an improvement of the local environmental quality, which is beneficial for seed germination and early survival. It therefore provides habitat for certain species that could not survive under the given conditions of the surroundings.
- we implement local facilitation as a positive effect on cell colonisation. 
- [equation] 
- Thus, the likelihood of 
- Thus, the positive effect is of great importance in arid environments.

Figure |

## local competition

- plants compete for resources like nutrients and water with other plants in their close vicinity. This effect leads to decellerated growth and eventually outcompetition of the weaker competitor on the local patch scale. Plants will benefit from competitor free space around their location, if the environment allows to grow there. 
- We implement local competition as a negative effect on growth. 
- [equation]
- This affects the local carrying capacity and therefore the likelihood of establishment of the system linearly with increasing density of plants in the surroundings.

Figure |

## associational resistance

- In rangelands, plants are subject to feeding pressure by livestock. In drylands, many plants have evolved defence mechanisms that protect them from being eaten, such as thorns or indigestible plant metabolites. Plants that grow next to each other share this investment and benefit from the structures provided by their neighbors. 
- We implement this as a reduction of plant mortality due to livestock feeding with increasing plant cover in the direct neighborhood. 
- [equation]
-

Figure |

## attractant decoy

- In contrast to the protective effects of plants onto each other,  accumulations of plants will also attract livestock and increase feeding pressure locally. 
- We implement this mechanism as a positive effect of local plant cover on the feeding rate of grazers, rendering more efficient. 
- [equation]

# Discussion

The different mechanisms alter the possible ecosystem states observed along a gradient of pressure, but they also alter the trajectories towards those attractors departing from weakly or highly clustered landscapes.

# Methods

We implement the equations for mortality and growth outlined above in a pair-approximation model of a spatially-explicit cellular automata framework. We contrast this to a model that does not account for spatial structure and assumes that the local neighborhood cover of plants is equal to the global vegetation cover, i.e. a mean field model.

## cellular automata model

## pair-approximation model

## mean-field model

```{r}
stopCluster(cl)
```

# References
